---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
#1. Defining the question

## 1.1 Specifying the data analytic objective

1. Perform clustering stating insights drawn from your analysis and visualizations.
2. Upon implementation, provide comparisons between the approaches learned this week i.e. K-Means clustering vs Hierarchical clustering highlighting the strengths and limitations of each approach in the context of your analysis.


## 1.2 Defining the metric of success

Clustering the data to maximize reaching the target audience


## 1.3 Understanding the context

Kira Plastinina is a Russian brand that is sold through a defunct chain of retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia. The brand’s Sales and Marketing team would like to understand their customer’s behavior from data that they have collected over the past year. More specifically, they would like to learn the characteristics of customer groups.

## 1.4 Recording the Experimental Design

1. Loading the data
2. Checking the data
3. Tidying the data
4. Univariate Analysis
5. Bivariate Analysis
6. Challenging the solution
7. Recommendations
8. Follow up questions

# 2. Loading the libraries

```{r}

# Installing packages that we have not.

#library(devtools)
#install_github("vqv/ggbiplot", force = TRUE)
#install.packages("DataExplorer") 
#install.packages("Hmisc")
#install.packages("pastecs")
#install.packages("psych")
#install.packages("corrplot")
#install.packages("factoextra")
#install.packages("Rtsne")
#install.packages("caret")

```

```{r}
# Loading Libraries necessary

library(tidyverse)
library(magrittr)
library(warn = -1)

library(ggbiplot)
library(RColorBrewer)
library(ggplot2)
library(lattice)
library(corrplot)

library(DataExplorer)
library(Hmisc)
library(pastecs)
library(psych)
library(factoextra)
library(Rtsne)
library(caret)
```

# 3. Loading Data 

```{r}

customers = read_csv('http://bit.ly/EcommerceCustomersDataset')
head(customers)
```

```{r}
tail(customers)
```

## structure of the data

```{r}
str(customers)
```
The data has 12330 obs. of  18 variables

## Checking for the summary description of our data

```{r}

# Checking for the summary description of our data

summary(customers)
```

# 4. DATA CLEANING

```{r}
colSums(is.na(customers))
```
We have missing values in 8 columns. Since we have quite a number of rows, we will go ahead and drop these missing values as we will be left with enough data for our analysis

```{R}
# creating a new data frame that does not have missing values

customers1 <- na.omit(customers)
head(customers1)
```

```{r}

# Confirming that we have no null values

sum(colSums(is.na(customers1)))
```

```{R}
# Checking for Duplicates

customer <- customers1[duplicated(customers1),]
dim(customer)
```

```{r}
# Removing these duplicated rows in the dataset
cust <- customers1[!duplicated(customers1), ]
dim(cust)
```

## Checking for ourliers

```{r}
# Administrative_Duration
# Plot a boxplot to help us visualise any existing outliers 
boxplot(cust$Administrative_Duration)
```

```{r}
# Informational_Duration
boxplot(cust$Informational_Duration)
```

```{r}
# ProductRelated_Duration
boxplot(cust$ProductRelated_Duration)
```

```{r}
# BounceRates
boxplot(cust$BounceRates)
```

```{r}
# PageValues
boxplot(cust$PageValues)
```

```{r}
# SpecialDay
boxplot(cust$SpecialDay)
```


We have outliers in several of our numerical columns. We shall not delete the outliers as they will result in us losing so much customer data which could alter our analysis. 

# 5. Exploratory Data Analysis

## Univariate Analysis

```{r}

# Plotting a histogram using ggplots 
# 
#

cust %>%
    ggplot(aes(ProductRelated)) +
    geom_histogram(color = "Green",fill = "yellow") +
    geom_vline(xintercept = mean(cust$ProductRelated), lwd = 2) +
    labs(title = "Distribution of Product Related",
         x = "Product Related",
         y = "Frequency")
```

```{r}

# plotting a histogram of Exit Rates

hist(cust$ExitRates,
     main = "Histogram of Exit Rates",
     xlab = "Exit Rates",
     col = "magenta")

```

```{r}
# Plotting all histograms in the continuous variables in our data 

plot_histogram(cust)
```

From the histograms, most of our variables are positively skewed.

```{r}
# Bar plots of the categorical/factor modes variables

#par(mfrow=c(4,1))

#for(i in 11:16) {
  
#	counts <- table(cust[,i])
	
#	name <- names(cust)[i]
#	barplot(counts, main=name, col = heat.colors(20))}
```


May and November were busy months receiving high traffic, Feb received the least traffic of customers.
Most vistors were returning type.
Traffic mode number 2, 1 and 3 were heavily used in that order.
Region number 1 had the most activity, region 5 was less active.
Browser 2 and 1 were the most commonly used for browsing.
Operating systems 2, 1 and 3 were mostly used by customers.


## Bivariate Analysis

```{r}
# Pair plots for the continous variables
pairs(cust[,1:9])
```



```{r}

# Correlations

# subseting our data frame to get the numeric variables
numerics <- cust[, c(1:10)]

# getting the correlation between these numeric variables
numerics.cor <- cor(numerics)
numerics.cor
```

```{r}
# installing packages that we shall use to plot the correlation plots

install.packages("Hmisc")
library("Hmisc")

install.packages("corrplot")
library(corrplot)

corrplot(numerics.cor)
```

```{r}
# Plotting a scatter plot

plot(ExitRates ~ BounceRates, data = cust, 
      col = "yellow",
      main = "Bounce vs Exit Rates Scatter Plot")
```


There is a strong positive correlation between bounce rates and exit rates

```{r}
plot(ProductRelated_Duration ~ ProductRelated, data = cust, 
      col = "yellow",
      main = "Product related vs Product related durations Scatter Plot")
```


The is a positive correlation between product relation and product relation duration

# K-MEAN CLUSTERING

```{r}
# converting  some of our columns into numerical data types by one hot encoding
dmy = dummyVars(" ~ .", data = cust)

df4 = data.frame(predict(dmy, newdata = cust))
```

```{r}

# Checking the data types of each attribute
sapply(df4, class)
```

```{r}

# Confirming changes
glimpse(df4)

```

```{r}
# We are instructed to use Revenue as the class label,
# Hence we will remove it and store it in another variable

df4_copy <- df4[, -c(30:31)]
cust.class<- cust[, "Revenue"]

df4_copy_copy <- df4[, -c(30,31)]
```

```{r}

# Previewing the class column

head (cust.class)
```

```{r}

# Previewing the copy dataset with dummies
head(df4_copy)
```

```{r}

# Normalizing the a copy of the original data

df_norm <- as.data.frame(apply(df4_copy, 2, function(x) (x - min(x))/(max(x)-min(x))))
```


```{r}
# Applying K-Means  Clustering algorithm 
# Using 3 centroids as K=3

result <- kmeans(df_norm, 10)
```

```{r}
# Previewing the number of records in each cluster

result$size
```

```{r}
# Viewing the cluster center data points by each attribute

result$centers
```

```{r}

# Plotting to see how exit rates and bounce rates data points have been distributed in clusters

plot(df_norm[c(7,8)], col = result$cluster)
```

```{r}
# Product Related, vs Product Related Duration

plot(df_norm[, 5:6], col = result$cluster)
```


# Hierachical Clustering

```{r}


# We use R function hclust() 
# For hierarchical clustering
# First we use the dist() to compute the Euclidean distance btwn obs
# d will be the first argument in the hclust() dissimilairty matrix
# 

d <- dist(df_norm, method = "euclidean")

# We then apply hierarchical clustering using the Ward's method

res.hc <- hclust(d, method = "ward.D2")

# Lastly we plot the obtained dendrogram
#--

plot(res.hc, cex = 0.6, hang = -1)

```




